---
title: "Scatterplots"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Read fire Incidents file
fire_Incidents <- readRDS("app/data/Trimmed_Fire_Incidents.Rds")
fire_Stations <- readRDS("app/data/Trimmed_Fire_Station.Rds")
neighbourhood_Shape <- readRDS("app/data/neighbourhoods_info.Rds")

library(ggplot2)
library(dplyr)
library(sf)
```

## Topics

* Average response time for incidents in the neighborhood (if takes longer means more stations should be built? Table? plot?)
* Average response time vs neighborhood income (if poor take longer than something shou change)
* Fire size vs number of personnel (if no trend means city should take more action to prioritize urgency)
* High Casualty(3+,5+?) vs Income or high casualty vs dwelling density (means the area need more equipment if poor; or means area needs more inspection if more building gets more casualty)
* Avg financial loss in each origin? (loss more ones would get discount from insurance if purchase fire equipment?)
* Which neighborhoods have severe fires (large, extra large etc but least station)? Which has the most fire (but least station)?
* Does a certain premise type area have more fire (commercial neighborhood, apartment, condo, houses etc)


# Top options

## Average response time

```{r Average response time vs neighborhood income}

#################################

# Group by date
fire_Incidents_Filtered <- fire_Incidents %>%
  group_by(Neighborhood) %>%
  summarise(Avaerge_Response_Time = mean(TFS_Response_Time), .groups = 'drop') %>%
  na.omit()

fire_Incidents_Filtered <- merge(fire_Incidents_Filtered, neighbourhood_Shape, by.y = "AREA_NAME", by.x="Neighborhood")

# Plot the scatter plot
plot_title <- "Average response time vs Usual Density"

ggplot(fire_Incidents_Filtered, aes(x=DWELLINGS_OCCUPIED_BY_USUAL_RESIDENTS/LAND_AREA_IN_SQKM, y=Avaerge_Response_Time)) + 
  geom_point()+ ggtitle(plot_title)+ coord_flip()+
  xlab("Usual Density")+ ylab("Avaerge Response Time")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

plot_title <- "Average response time vs Average Household Sizes"

ggplot(fire_Incidents_Filtered, aes(x=AVERAGE_HOUSRHOD_SIZE, y=Avaerge_Response_Time)) + 
  geom_point()+ ggtitle(plot_title)+ coord_flip()+
  xlab("Average households ")+ ylab("Avaerge Response Time")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

```
## Casualty

```{r High Casualty(3+,5+?) vs Income or high casualty vs dwelling density}

# Group by date
fire_Incidents_Filtered <- fire_Incidents %>%
  group_by(Neighborhood) %>%
  summarise(Avaerge_Casualty = mean(Civilian_Casualties), .groups = 'drop') %>%
  na.omit()

fire_Incidents_Filtered <- merge(fire_Incidents_Filtered, neighbourhood_Shape, by.y = "AREA_NAME", by.x="Neighborhood")

# Plot the scatter plot
plot_title <- "Average Casualty vs Population Density"

ggplot(fire_Incidents_Filtered, aes(x=Avaerge_Casualty, y=POPULATION_DENSITY_PER_SQKM)) + 
  geom_point()+ ggtitle(plot_title)+
  xlab("Average Casualty")+ ylab("Population Density")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

# Plot the scatter plot
plot_title <- "Average Casualty vs Dwellings Density by Neibourghood Sizes"

ggplot(fire_Incidents_Filtered, aes(x=Avaerge_Casualty, y=DWELLINGS_OCCUPIED_BY_USUAL_RESIDENTS/LAND_AREA_IN_SQKM)) + 
  geom_point()+ ggtitle(plot_title)+geom_smooth(stat = 'smooth',method = 'loess', formula = y ~ x)+
  xlab("Average Casualty")+ ylab("Dwellings Density by Neibourghood Sizes")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

# Plot the scatter plot
plot_title <- "Average Casualty vs Percentage Residents with No Certificate"

ggplot(fire_Incidents_Filtered, aes(x=Avaerge_Casualty, y=PERCENTAGE_NO_CERTIFICATE_DIPLOMA)) + 
  geom_point()+ ggtitle(plot_title)+geom_smooth(stat = 'smooth',method = 'loess', formula = y ~ x)+
  xlab("Average Casualty")+ ylab("Percentage Residents with No Certificate")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

```
## Avg financial loss

```{r Avg financial loss in each origin?}

# Group by date
fire_Incidents_Filtered <- fire_Incidents %>%
  group_by(Neighborhood) %>%
  summarise(Average_Financial_Loss = mean(Estimated_Dollar_Loss), .groups = 'drop') %>%
  na.omit()

fire_Incidents_Filtered <- merge(fire_Incidents_Filtered, neighbourhood_Shape, by.y = "AREA_NAME", by.x="Neighborhood")

# Plot the scatter plot
plot_title <- "Average Financial Loss vs Average Household Income"

ggplot(fire_Incidents_Filtered, aes(y=AVERAGE_HOUSEHOLD_INCOME, x=Average_Financial_Loss)) + 
  geom_point()+ ggtitle(plot_title)+ geom_smooth(stat = 'smooth',method = 'loess', formula = y ~ x)+
  xlab("Average_Financial_Loss")+ ylab("AVERAGE_HOUSEHOLD_INCOME")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

# Plot the scatter plot
plot_title <- "Average Financial Loss vs Average Household Size"

ggplot(fire_Incidents_Filtered, aes(y=AVERAGE_HOUSRHOD_SIZE, x=Average_Financial_Loss)) + 
  geom_point()+ ggtitle(plot_title)+ geom_smooth(stat = 'smooth',method = 'loess', formula = y ~ x)+
  xlab("Average Financial Loss")+ ylab("Average Household Size")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))
```
## Total Incidents

```{r Which neighborhoods have severe fires (large, extra large etc but least station)? Which has the most fire (but least station)?}


plot_title <- "Total Incidents vs Total Stations"

ggplot(neighbourhood_Shape, aes(x=TOTAL_STATION, y=TOTAL_INCIDENTS)) + 
  geom_point()+ ggtitle(plot_title)+ coord_flip()+
  xlab("TOTAL_STATION")+ ylab("TOTAL_INCIDENTS")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

plot_title <- "Total Incidents vs Population Density"

ggplot(neighbourhood_Shape, aes(x=POPULATION_DENSITY_PER_SQKM, y=TOTAL_INCIDENTS)) + 
  geom_point()+ ggtitle(plot_title)+ coord_flip()+
  xlab("Population Density")+ ylab("TOTAL_INCIDENTS")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

```
# Other Options

```{r Fire size vs number of personnel}

# Plot the scatter plot
plot_title <- "Fire size vs number of personnel"

ggplot(fire_Incidents, aes(x=Fire_Size_Case, y=Number_of_responding_personnel)) + 
  geom_bar(stat='identity')+ ggtitle(plot_title)+ coord_flip()+
  xlab("Fire Size")+ ylab("Number of responding personnel")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

```

```{r Average response time for incidents in the neighborhood}

# Group by date
fire_Incidents_Filtered <- fire_Incidents %>%
  group_by(Neighborhood) %>%
  summarise(Avaerge_Response_Time = mean(TFS_Response_Time), .groups = 'drop') %>%
  na.omit()

fire_Incidents_Filtered <- fire_Incidents_Filtered[which(order(fire_Incidents_Filtered$Avaerge_Response_Time) < 10), ]

# Plot the scatter plot
plot_title <- "Average response time for incidents top 40"

ggplot(fire_Incidents_Filtered, aes(x=Neighborhood, y=Avaerge_Response_Time)) + 
  geom_bar(stat='identity')+ ggtitle(plot_title)+ coord_flip()+
  xlab("Neighborhood")+ ylab("Avaerge Response Time")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))

fire_Incidents_Filtered

```

```{r Does a certain premise type area have more fire}

plot_title <- "Does a certain premise type area have more fire?"

ggplot(fire_Incidents, aes(x=Area_Orgin_Case)) + 
  geom_bar()+ ggtitle(plot_title)+ coord_flip()+
  xlab("Area_Orgin_Case")+ ylab("TOTAL_INCIDENTS")+
  theme(plot.title = element_text(color="Black", size=14, face="bold"))
```